Transitivity of Subtyping for Intersection Types

### 2020-05-14 papers/intersection-transitivity.tex
check

\begin{abstract}
  The subtyping rules for intersection type systems traditionally
  employ a transitivity rule (Barendregt et al. 1983), which means
  that subtyping does not enjoy the subformula property.  Laurent
  develops a sequent-style subtyping system, without transitivity, and
  proves transitivity via a sequence of six lemmas that culminate in
  cut-elimination (2018). This article develops a subtyping system in
  regular style that omits transitivity and provides a direct proof of
  transitivity. The new system satisfies something like the subformula
  property: every type occuring in the subtyping derivation of $A <:
  B$ is a subformula of $A$ or $B$, or an intersection of such
  subformulas.  Borrowing from Laurent's system, the rule for function
  types is essentially the $\beta$-soundness property.  The main lemma
  required for the transitivity proof is one that has been used to
  prove the inversion principle for subtyping of function types. The
  choice of induction principle for the proof of transitivity is
  subtle: we use well-founded induction on the lexicographical
  ordering of the sum of the depths of the first and last type
  followed by the sum of the sizes of the middle and last type. The
  article concludes with a proof that the new subtyping judgment is
  equivalent to that of Barendregt, Coppo, and Dezani-Ciancaglini.
\end{abstract}

@@

Intersection types were invented by Coppo, Dezani-Ciancaglini, and
Salle, as a tool for studying normalization in the lambda
calculus~\citep{Coppo:1979aa}. Subsequently intersection types have
been used for at least three purposes: to enhance the expressiveness
of type systems
~\citep{Reynolds:1988aa,Pierce:1991aa,Castagna:2014aa,Chaudhuri:2014aa,Oliveira:2016aa,Muehlboeck:2018aa,Bi:2019aa,Dunfield:2019aa,Microsoft:TypeScript2020aa,Dotty:2020aa},
to increase the precision of static
analyses~\citep{Turbak:1997aa,Palsberg:1998aa,Mossin:2003aa,Simoes:2007aa},
and to express filter models for the denotational semantics of a
variety of lambda calculi
~\citep{Coppo:1980ab,Coppo:1981aa,Engeler:1981aa,Coppo:1984aa,Honsell:1992aa,Abramsky:1993fk,Plotkin:1993ab,Honsell:1999aa,Ishihara:2002aa,Rocca:2004aa,Dezani-Ciancaglini:2005aa,Alessi:2006aa}.

UNDER CONSTRUCTION

>>>>

\begin{abstract}
  The subtyping rules for intersection type systems traditionally
  employ a transitivity rule (Barendregt et al. 1983), which means
  that subtyping does not enjoy the subformula property.  Laurent
  develops a sequent-style subtyping system, without transitivity, and
  proves transitivity via a sequence of six lemmas that culminate in
  cut-elimination (2018). This article develops a subtyping system in
  regular style that omits transitivity and provides a direct proof of
  transitivity, significantly reducing the length of the proof. The
  new system satisfies something like the subformula property: every
  type occuring in the derivation of $A <: B$ is a subformula of $A$
  or $B$, or an intersection of such subformulas.  Borrowing from
  Laurent's system, the rule for function types is essentially the
  $\beta$-soundness property.  The main lemma required for the
  transitivity proof is one that has been used to prove the inversion
  principle for subtyping of function types. The article concludes
  with a proof that the new subtyping system is equivalent to that of
  Barendregt, Coppo, and Dezani-Ciancaglini.
\end{abstract}

@@

Intersection types were invented by Coppo, Dezani-Ciancaglini, and
Salle as a tool for studying normalization in the lambda
calculus~\citep{Coppo:1979aa}. Subsequently intersection types have
been used for at least three purposes: to enhance the expressiveness
of type systems
~\citep{Reynolds:1988aa,Pierce:1991aa,Castagna:2014aa,Chaudhuri:2014aa,Oliveira:2016aa,Muehlboeck:2018aa,Bi:2019aa,Dunfield:2019aa,Microsoft:TypeScript2020aa,Dotty:2020aa},
to increase the precision of static
analyses~\citep{Turbak:1997aa,Palsberg:1998aa,Mossin:2003aa,Simoes:2007aa},
and to express filter models for the denotational semantics of a
variety of lambda calculi
~\citep{Coppo:1980ab,Coppo:1981aa,Engeler:1981aa,Coppo:1984aa,Honsell:1992aa,Abramsky:1993fk,Plotkin:1993ab,Honsell:1999aa,Ishihara:2002aa,Rocca:2004aa,Dezani-Ciancaglini:2005aa,Alessi:2006aa}.

The work in this
